---
title: "Spatial Optimization for Pysolo"
author: "Erico Kutchartt, Francesco Pirotti"
date: "2026-04-26"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(terra)
library(sf)
library(tidyverse)
library(spopt)
library(tidyverse)
library(mapgl)
library(viridisLite)
library(htmlwidgets)
library(dplyr)
## biomass suitability color scale
val_range <- c(0, 0.1, 0.2, 0.3)
v_cols <- viridisLite::viridis(4)

nuts <- sf::read_sf("/archivio/shared/geodati/vector/NUTS_2024_all_4326v2.gpkg")
source("pocForestaleFunctions.R")
sinkf <- file("output/benchmark.csv", "a")
if(!file.exists("pts.rda")){
  stop("Please run pocForestalePrep.R first")
}
if(!exists("pts")) {
  load(file="pts.rda")
}
if(!file.exists("data/cities3.gpkg")){
  stop("Please run pocForestalePrep.R first")
}

cities <- sf::read_sf("data/citiesFinal.gpkg") #|> select(3) #|> sf::st_transform(3035)
df <- list()
ccode <- list( "Greece"="EL", "Italy"="IT", "Spain"="ES")
cname <- names(ccode)
names(cname) <- ccode
countries <- cname[ unique(cities$CNTR_CODE) ]


```

## Prepare data

This is a classic large-scale Facility Location Problem (FLP), specifically a variation of the Capacitated Plant Location Problem with a "winner-take-all" constraint on the biomass points.

### The "Greedy" Heuristic (Finding the "Ideal Number")

To find the ideal number of cities without running complex solvers:Calculate Marginal Utility: For each city, calculate its potential profit if it took all nearby unclaimed biomass.Rank and Pick: Pick the city with the highest profit. Mark its biomass points as "taken."Iterate: Repeat until the next best city's Gain ($Y$) minus Loss ($X$) and transport costs becomes negative.The Stop Point: The moment your next best city adds more cost than it brings in biomass value, you have found your Ideal Number.

```{r processit, echo=FALSE, results="asis"}


  cat("Country\tN cities\tTime of processing\n\n")

for(countryn in names(ccode)){
  # dirout <- sprintf("%s/%s", out_dir, countryn)
  # if(!dir.exists(dirout)){
  #   dir.create(dirout)
  # }
  cat("## ", countryn, "\t\n\n")
  candidateCitiest <- sf::read_sf(sprintf("%s_cityScore3.gpkg", countryn))

  
  candidateCities <- candidateCitiest |> filter(dni > 1400)  
  cat("### ", nrow(candidateCities), "/",nrow(candidateCitiest), " candidate cities after DNI filter at 1400\n\n")
 
  for(i in 1:3){
      vv <- cbind(sf::st_coordinates(candidateCities)/10, 
              candidateCities[[sprintf("sumBiomassSuitabilityT%dn", i)]])
      km <- kmeans(vv, 10)
      candidateCities$k <- km$cluster
      
      cluster_bounds <- candidateCities %>%
        group_by(k) %>%
        summarize(geometry = st_union(geom)) %>% # Merge points in same cluster
        st_convex_hull()                             # Draw the boundary around them
      
      suppressWarnings(plotCluster(candidateCities, cluster_bounds, ret=T, ccode[[countryn]]) )
  }


}

close(sinkf)
```
 
